\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-01-02 p2/p2.Rmd
.

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  all <- all %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all, file="../data/all-profiles.rds")
} else {
  all <- readRDS("../data/all-profiles.rds")
}

errors2 <- rbind(errorrate(all, 0.001), 
                errorrate(all, 0.005), 
                errorrate(all, 0.01), 
                errorrate(all, 0.05))
```

@@

window Optimization 320 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==320, wv==50) %>% xtabs(data=., ~signif+match)
```
window Optimization 120 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==120, wv==50) %>% xtabs(data=., ~signif+match)
```
window Optimization 80 Window Validation 50
```{r echo=FALSE}
all %>% filter(wo==80, wv==50) %>% xtabs(data=., ~signif+match)

@@

```{r, echo=FALSE, results='asis'}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
```

@@

```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
ns <- all %>% group_by(wv, wo, match) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)
ns %>% filter(wv %in% c(30, 50)) %>% 
  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
  geom_smooth(method="lm", size=.5, se=FALSE) +
  geom_point() + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Size of optimization window, wo") +
  scale_colour_brewer("Match", palette="Set1") +
  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
```
Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e. tests in which a particular parameter setting did not return a valid result. This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation. The problem is therefore exacerbated by a larger validation window. 
Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.

@@

```{r}
errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +
  geom_point(data=errors2, aes(shape="profiles"))
```
## Discussion

>>>>

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  allp <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    allp <- rbind(allp, tmp)
  }
  allp <- allp %>% filter(land1_id < land2_id)
  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(allp, file="../data/all-profiles.rds")
} else {
  allp <- readRDS("../data/all-profiles.rds")
}

errors2 <- rbind(errorrate(allp, 0.001), 
                errorrate(allp, 0.005), 
                errorrate(allp, 0.01), 
                errorrate(allp, 0.05))

```{r data-sig-coars}
if (!file.exists("../data/all-sig-c.rds")) {
  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
  all_sig_c <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all_sig_c <- rbind(all_sig_c, tmp)
  }
  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
} else {
  all_sig_c <- readRDS("../data/all-sig-c.rds")
}

errors_sig_c <- rbind(errorrate(all_sig_c, 0.001), 
                errorrate(all_sig_c, 0.005), 
                errorrate(all_sig_c, 0.01), 
                errorrate(all_sig_c, 0.05))

@@

<!-- \begin{minipage}{0.3\textwidth} -->
```{r }
library(kableExtra)
a27<- all %>% filter(wo==320, wv==50) %>% xtabs(data= ., ~signif+match)
knitr::kable(a27, digits=0, caption = "Window of Optimization 320 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))

```
<!-- \end{minipage} -->
<!-- \begin{minipage}{0.3\textwidth} -->
```{r}
a27<- all %>% filter(wo==120, wv==50) %>% xtabs(data= ., ~signif+match)
knitr::kable(a27, digits=0, caption = "Window of Optimization 120 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
```
<!-- \end{minipage} -->
```{r}
a27<- all %>% filter(wo==80, wv==50) %>% xtabs(data= ., ~signif+match)
knitr::kable(a27, digits=0, caption = "Window of Optimization 80 and Window of Validation 50", format = "latex" ,booktabs=TRUE)%>% kableExtra::kable_styling(latex_options = c("hold"))
```

@@

```{r, echo=FALSE, results='asis'}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
knitr::kable(t(param_setting)[,1:17], digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down", "hold"))

knitr::kable(t(param_setting)[,17:33], digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down","hold"))

# library(xtable)
# strCaption <- "Overview of parameter settings used for optimization and validation windows for bullet land signatures."
#   
#   # set up xtable output
# print(xtable(t(param_setting)[,1:17], 
#              caption = strCaption, label = ""),
#       include.rownames = TRUE,
#       include.colnames = FALSE,
#       caption.placement = "top",
#       comment=FALSE,
#       floating=FALSE
#       )
# print(xtable(t(param_setting)[,17:33],label = ""),
#       include.rownames = FALSE,
#       include.colnames = FALSE,
#       caption.placement = "bottom",
#       comment=FALSE,
#       floating=TRUE
#       #table.placement = "!ht"
#       )
```

@@

```{r missings, fig.width=8, fig.height=5, fig.cap="Number of failed tests by the window optimization size, wo, and ground truth. Test results from different sources have a much higher chance to fail, raising the question, whether failed tests should be treated as rejections of the null hypothesis of same source."}
ns <- all %>% group_by(wv, wo, match) %>% summarize(
  n_ss = mean(same_shift_n),
  n_ds = mean(diff_shift_n),
  zero_ss = sum(same_shift_n==0),
  zero_ds = sum(diff_shift_n==0),
  miss = sum(is.na(p_value)),
  missperc = miss/85491*100
)

#ns1<- ns[which(ns$wv == c(30,50), arr.ind = T),]
ns %>% filter(wv %in% c(30, 50)) %>%
#ns %>% filter(wv == c(30, 50)) %>%
  ggplot(aes(x = wo, y = missperc, shape=factor(wv), colour = factor(match))) + 
  geom_smooth(method="lm", size=.5, se=FALSE) +
  geom_point() + 
  theme_bw() +
  ylab("Percent of failed tests.") +
  xlab("Size of optimization window, wo") +
  scale_colour_brewer("Match", palette="Set1") +
  scale_shape_discrete("Size of \nvalidation\nwindow, wv")
```
Figure \ref{fig:missings} gives an overview of the number of failed tests, i.e. tests in which a particular parameter setting did not return a valid result. This happens, when the shift to align two signatures is so large, that the remaining overlap is too small to accommodate  windows for validation. The problem is therefore exacerbated by a larger validation window. 
Figure \ref{fig:missings} also shows that the number of failed tests is approximately linear in the size of the optimization window.

@@

```{r}
#errors %>% ggplot(aes(x = wv, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + facet_grid(~wo) +geom_point(data=errors2, aes(shape="profiles"))

errors_sig_c %>% ggplot(aes(x = coarse, y=total, colour=factor(alpha), shape="signatures")) + geom_point() + geom_line() + geom_point(data=errors2, aes(shape="profiles")) +geom_line(data=errors2, aes(shape="profiles"))
```
## Discussion

### 2018-01-02 p2/p2.tex
.

\section{Introduction}\label{introduction}

@@

\subsection{Existing solution for toolmarks: Chumbley
Score}\label{existing-solution-for-toolmarks-chumbley-score}

\hh{XXX anything to do with motivation has to come first - move this to the first paragraph of the intro XXX}
\hh{XXX data and results come after the intro to chumbley.}
\hh{XXX start with defining the data, i.e. we are working with random processes $x(t)$ and $y(t)$ ... you should also go into the coarseness parameter here - that is a pre-processing step before the optimization. }
\gk{done}

@@

\subsection{Chumbley Algorithm and Chumbley
score}\label{chumbley-algorithm-and-chumbley-score}

@@

\subsection{\texorpdfstring{Potential limitations in the application to
striation marks on bullet lands: smaller in width and curved - need to
adjust parameter settings; something similar is done in
\citet{afte-chumbley} for toolmark comparisons of slip-joint
pliers.}{Potential limitations in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in  for toolmark comparisons of slip-joint pliers.}}\label{potential-limitations-in-the-application-to-striation-marks-on-bullet-lands-smaller-in-width-and-curved---need-to-adjust-parameter-settings-something-similar-is-done-in-for-toolmark-comparisons-of-slip-joint-pliers.}

@@

\subsection{Using cross-validation setup to identify appropriate
parameter settings for (a) signatures and (b) profiles
directly}\label{using-cross-validation-setup-to-identify-appropriate-parameter-settings-for-a-signatures-and-b-profiles-directly}

Following on similar lines to the setup of toolamrks, the first step
here is to first identify what difference does different window sizes of
optimization and the validation step have, when adapting the toolmark
method to bullets.

@@

window Optimization 320 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 77564   412
  TRUE   4022   674
\end{verbatim}

window Optimization 120 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 79249   386
  TRUE   4523   854
\end{verbatim}

window Optimization 80 Window Validation 50

\begin{verbatim}
       match
signif  FALSE  TRUE
  FALSE 78817   450
  TRUE   4680   775
\end{verbatim}

@@

\caption{\label{tab:unnamed-chunk-5}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrrr}
\toprule
wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
\bottomrule
\end{tabular}}
\end{table}

@@

The idea therefore is to first use these raw values of the profile
directly in the chumbley algorithm, and see how the algorithm performs
for different coarseness values (smoothing parameter as referred in the
function LOWESS used in the chumbley algorithm)

>>>>

\tableofcontents

\section{Introduction and Background}\label{introduction-and-background}

@@

\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}

\hh{You need to distinguish between $x(t)$ and $x(t_1)$ - $x(t)$ is a spatial process for some $t$. $x(t_1)$ is the realization of the process in $t_1$. - Alternatively you could also follow the notation in Hadler et al on 'digitized toolmarks at a pixel level' - where our 'pixel' is the resolution of the confocal light microscope, i.e. a pixel corresponds to 0.645 microns.}

@@

\subsection{Detailed algorithm}\label{detailed-algorithm}

@@

\subsection{Potential limitations of the Chumbley
Score}\label{potential-limitations-of-the-chumbley-score}

\hh{in the application to striation marks on bullet lands: smaller in width and curved - need to adjust parameter settings; something similar is done in \citet{afte-chumbley} for toolmark comparisons of slip-joint pliers. XXX work in }

@@

\subsection{Testing setup:}\label{testing-setup}

\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}

Following on similar lines to the setup of toolmarks, the first step
here is to first identify what difference does different window sizes of
optimization and the validation step have, when adapting the toolmark
method to bullets.

@@

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-2}Window of Optimization 320 and Window of Validation 50}
\centering
\begin{tabular}[t]{lrr}
\toprule
  & FALSE & TRUE\\
\midrule
FALSE & 77564 & 412\\
TRUE & 4022 & 674\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Window of Optimization 120 and Window of Validation 50}
\centering
\begin{tabular}[t]{lrr}
\toprule
  & FALSE & TRUE\\
\midrule
FALSE & 78938 & 371\\
TRUE & 4281 & 845\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-4}Window of Optimization 80 and Window of Validation 50}
\centering
\begin{tabular}[t]{lrr}
\toprule
  & FALSE & TRUE\\
\midrule
FALSE & 78817 & 450\\
TRUE & 4680 & 775\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-5}Overview of parameter settings used for optimization and validation windows for bullet land signatures.}
\centering
\resizebox{\linewidth}{!}{\begin{tabular}[t]{lrrrrrrrrrrrrrrrrr}
\toprule
wo & 50 & 50 & 60 & 60 & 80 & 80 & 90 & 90 & 100 & 100 & 110 & 110 & 120 & 120 & 120 & 120 & 120\\
wv & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 10 & 20 & 30 & 50 & 60\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{\begin{tabular}{lrrrrrrrrrrrrrrrrr}
\toprule
wo & 120 & 130 & 130 & 140 & 140 & 150 & 150 & 160 & 160 & 200 & 200 & 200 & 240 & 240 & 280 & 280 & 320\\
wv & 60 & 30 & 50 & 30 & 50 & 30 & 50 & 30 & 50 & 20 & 30 & 50 & 30 & 50 & 30 & 50 & 50\\
\bottomrule
\end{tabular}}
\end{table}

@@

\hh{XXX ideas should go to the previous section on setting up the study; in this section we focus on results only XXX}
The idea therefore is to first use these raw values of the profile
directly in the chumbley algorithm, and see how the algorithm performs
for different coarseness values (smoothing parameter as referred in the
function LOWESS used in the chumbley algorithm)

{\centering \includegraphics[width=\textwidth]{figures/unnamed-chunk-6-1}

\caption{Type II errors for different levels of coarseness. XXX We need more values around .25}\label{fig:unnamed-chunk-6}
\end{figure}

\subsubsection{comparison of total error with signatures and
profiles}\label{comparison-of-total-error-with-signatures-and-profiles}

\begin{center}\includegraphics[width=\textwidth]{figures/unnamed-chunk-7-1} \end{center}

\subsection{Discussion}\label{discussion}

\subsection{Results of Chumbley score}\label{results-of-chumbley-score}

\begin{itemize}
\tightlist
\item
  Nominal alpha value shows dependency on the size of the window of
  optimization
\item
  Test Fail depends on whether known-match or known non-matches has
  predictive value
\item
  Type II error least bad for window of validation 30 and window of
  optimization 120
\end{itemize}

\section{Algorithm Modification}\label{algorithm-modification}

The proposed algorithm modification is at the same shift step which
comes after the shift distance is identified by finding two windows in
the two markings that have the highest correlation.

The modification allows for a ``wiggle'' room in the second marking of
the two sets of markings being compared. This means that each set of
same shift windows that are being compared, the second marking will
window that is under consideration is allowd to move a little towards
the left and a little towards the right.

This gives us a new set of 4 windows, 2 to the left and 2 to the right
of the original comparison window. Then the correlations for each one of
these 5 windows with the window under consideration of the 1st marking
is retrieved (from the validation step correlation matrix) or computed.

Then the window that has the maximum correlation (from the set of 5
windows in the 2nd marking) with the window of the 1st marking is
chosen.

This new maximum correlation is then used to compute the U-statistics
using the same method as before.

\subsection{Expected advantage of
modification}\label{expected-advantage-of-modification}

We are expecting that this modification would improve the type I and II
errors for the good. Less number of false negatives and false positives.
The reason for this expectation is many a times the bullet markings are
not made at the same distances for two or more bullets because of the
way it comes out of the barrel.

Therefore rigid same shifts might not necessarily compare the right set
of windows. In the same shift step for one set of windows, allowing the
window in the 2nd marking to wiggle left and right and finding the best
match to the window in the 1st marking, lets us adjust for situations
where markings are compressed or elongated.

\subsection{Dependence on the delta of new windows from the original
same shift
window}\label{dependence-on-the-delta-of-new-windows-from-the-original-same-shift-window}

From initial tests the amount of movement to get new windows left and
right of the original same shift window seems to directly influence the
the pvalue and U statistic that we look at.

\subsubsection{other questions}\label{other-questions}

How much movement is reasonable and should be used? Is there a way to
understand if the 2nd marking in comparison to the 1st marking is
compressed or elongated or neither? Should the amount of movement
selected depend on this compression or elongation? or using a static
delta movement value justifieable.
